---
output:
  word_document: default
  html_document: default
---
# Final Project: Phase 1
## MIS 502
## 06/13/2026


# Phase 1 Project description
For Phase 1 you will conduct a thorough exploratory/descriptive analysis of the dataset. Please **DO NOT** build any predictive models (e.g., logistic regression, trees, etc.) in this phase. Assume that your "audience" for this work are non-technical decision-makers.   




```{r loading libraries, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(GGally)
library(ggcorrplot)
library(gridExtra)
library(mice)
library(VIM)
library(naniar)
library(skimr)
library(UpSetR)
library(knitr)
library(kableExtra)
```





# Bringing in data set
```{r}
train <- read_csv("train.csv")

train = train %>% mutate_if(is.character, as.factor) %>% mutate (attribute_2 = as_factor(attribute_2)) %>% mutate (attribute_3 = as_factor(attribute_3)) %>% select(-id)

str(train)

```





# Looking at missing values
```{r}
skim(train) #lots of data missing
gg_miss_case(train) 
gg_miss_fct(x=train, fct=failure) #measurements 14-17 high number of missing data
gg_miss_upset(train, nsets=24) #doesn't seem like there's variables that are always missing together

#how many is lost if remove all NAs
lost_train = train %>% drop_na()
nrow(train) - nrow(lost_train)

#would lose 14387 variables - over half the dataset - will want to impute
```






# Looking at individual variables of interest
```{r}
ggplot(train, aes(x=measurement_0)) + labs(title="Distribution of Measurement 0", x = "Measurement 0", y="Count") + geom_histogram() + theme_bw()
ggplot(train, aes(x=measurement_1)) + geom_histogram() + labs(title="Distribution of Measurement 1", x = "Measurement 1", y="Count") + geom_histogram() + theme_bw()
ggplot(train, aes(x=measurement_2)) + geom_histogram() + labs(title="Distribution of Measurement 2", x = "Measurement 2", y="Count") + geom_histogram() + theme_bw()
ggplot(train, aes(x=measurement_3)) + geom_histogram()
ggplot(train, aes(x=measurement_4)) + geom_histogram()
ggplot(train, aes(x=measurement_5)) + geom_histogram()
ggplot(train, aes(x=measurement_6)) + geom_histogram()
ggplot(train, aes(x=measurement_7)) + geom_histogram()
ggplot(train, aes(x=measurement_8)) + geom_histogram()
ggplot(train, aes(x=measurement_9)) + geom_histogram()
ggplot(train, aes(x=measurement_10)) + geom_histogram()
ggplot(train, aes(x=measurement_11)) + geom_histogram()
ggplot(train, aes(x=measurement_12)) + geom_histogram()
ggplot(train, aes(x=measurement_13)) + geom_histogram()
ggplot(train, aes(x=measurement_14)) + geom_histogram()
ggplot(train, aes(x=measurement_15)) + geom_histogram()
ggplot(train, aes(x=measurement_16)) + geom_histogram()
ggplot(train, aes(x=measurement_17)) + geom_histogram()
```





# Visualizing variable relationships to failure
```{r}
product_code = ggplot(train, aes (x = product_code, fill=failure)) + geom_bar(position = "fill")
loading = ggplot(train, aes (x = loading, y=failure)) + geom_boxplot()
attribute_0 = ggplot(train, aes (x = attribute_0, fill=failure)) + geom_bar(position = "fill")
attribute_1 = ggplot(train, aes (x = attribute_1, fill=failure)) + geom_bar(position = "fill")
attribute_2 = ggplot(train, aes (x = attribute_2, fill=failure)) + geom_bar(position = "fill")
attribute_3 = ggplot(train, aes (x = attribute_3, fill=failure)) + geom_bar(position = "fill")
measurement_0 = ggplot(train, aes (x = measurement_0, y=failure)) + geom_boxplot()
measurement_1 = ggplot(train, aes (x = measurement_1, y=failure)) + geom_boxplot()
measurement_2 = ggplot(train, aes (x = measurement_2, y=failure)) + geom_boxplot()
measurement_3 = ggplot(train, aes (x = measurement_3, y=failure)) + geom_boxplot()
measurement_4 = ggplot(train, aes (x = measurement_4, y=failure)) + geom_boxplot()
measurement_5 = ggplot(train, aes (x = measurement_5, y=failure)) + geom_boxplot()
measurement_6 = ggplot(train, aes (x = measurement_6, y=failure)) + geom_boxplot()
measurement_7 = ggplot(train, aes (x = measurement_7, y=failure)) + geom_boxplot()
measurement_8 = ggplot(train, aes (x = measurement_8, y=failure)) + geom_boxplot()
measurement_9 = ggplot(train, aes (x = measurement_9, y=failure)) + geom_boxplot()
measurement_10 = ggplot(train, aes (x = measurement_10, y=failure)) + geom_boxplot()
measurement_11 = ggplot(train, aes (x = measurement_11, y=failure)) + geom_boxplot()
measurement_12 = ggplot(train, aes (x = measurement_12, y=failure)) + geom_boxplot()
measurement_13 = ggplot(train, aes (x = measurement_13, y=failure)) + geom_boxplot()
measurement_14 = ggplot(train, aes (x = measurement_14, y=failure)) + geom_boxplot()
measurement_15 = ggplot(train, aes (x = measurement_15, y=failure)) + geom_boxplot()
measurement_16 = ggplot(train, aes (x = measurement_16, y=failure)) + geom_boxplot()
measurement_17 = ggplot(train, aes (x = measurement_17, y=failure)) + geom_boxplot()


grid.arrange(product_code, loading)
grid.arrange(attribute_0, attribute_1, attribute_2, attribute_3)
grid.arrange(measurement_0, measurement_1, measurement_2, measurement_3)
grid.arrange(measurement_4, measurement_5, measurement_6, measurement_7)
grid.arrange(measurement_8, measurement_9, measurement_10, measurement_11)
grid.arrange(measurement_12, measurement_13, measurement_14, measurement_15)
grid.arrange(measurement_16, measurement_17)

grid.arrange(product_code, loading, attribute_0, attribute_1, attribute_3)
grid.arrange(measurement_0, measurement_1, measurement_2, measurement_3, measurement_4, measurement_5, measurement_6, measurement_7, measurement_8, measurement_9, measurement_10, measurement_11, measurement_12, measurement_13, measurement_14, measurement_15, measurement_16, measurement_17)
```





# Formatting potentially important variables
```{r}
loading = ggplot(train, aes (x = loading, y=failure)) + geom_boxplot() + labs(title = "Relationship between loading and product failure", x="Loading", y="Failure") + theme_bw()
loading

measurement_17 = ggplot(train, aes (x = measurement_17, y=failure)) + geom_boxplot() + labs(title = "Relationship between measurement 17 and product failure", x="Measurement 17", y="Failure") + theme_bw()
measurement_17

attribute_2 = ggplot(train, aes (x = attribute_2, fill=failure)) + labs(title= "Failure rates by attribute 2", y= "Proportion of failure results", x= "Attribute 2") + geom_bar(position = "fill") + theme_bw()
attribute_2

product_code = ggplot(train, aes (x = product_code, fill=failure)) + labs(title= "Failure rates by product code", y= "Proportion of failure results", x= "Product Code") + geom_bar(position = "fill") + theme_bw()
product_code

```





# Looking at correlations between all variables
```{r}
ggcorr(train, size=2)   

#Looks like measurement 17 may be correlated with measurement 4-9, taking a closer look

ggpairs(train, columns = c("failure", "measurement_4", "measurement_5", "measurement_6", "measurement_7", "measurement_8", "measurement_9", "measurement_17"))

#measurement 5 and 8 strongest correlation with measurement 17


```





# Looking deeper at correlations
```{r}
measurement_5x17 = ggplot(train, aes (x = measurement_5, y=measurement_17)) + labs(title = "Measurement 5 and Measurement 17 Correlation", x = "Measurement 5", y="Measurement 17") + geom_point() + geom_smooth(method = "lm") + theme_bw() 
measurement_5x17
#as measurement 5 increases, so does measurement 17

measurement_8x17 = ggplot(train, aes (x = measurement_8, y=measurement_17)) + labs(title = "Measurement 8 and Measurement 17 Correlation", x = "Measurement 8", y="Measurement 17") + geom_point() + geom_smooth(method = "lm") + theme_bw() 
measurement_8x17
#as measurement 5 increases, so does measurement 17

```





# Is there an interaction effect?
```{r}
recipe = recipe(failure ~ measurement_5 + measurement_8 + measurement_17, train)

model = logistic_reg( mode="classification") %>%
  set_engine("glm")


workflow = 
  workflow() %>% 
  add_model(model) %>% 
  add_recipe(recipe)

lm_fit = fit(workflow, train)
summary(lm_fit$fit$fit$fit)

car::vif(lm_fit$fit$fit$fit)

#VIF < 4: likely don't need to be worried about interaction effects

```





# Closer look at the attributes
```{r}
attribute_0x1 = ggplot(train, aes (x = attribute_0, fill=attribute_1)) + geom_bar()
attribute_0x2 = ggplot(train, aes (x = attribute_0, fill=attribute_2)) + geom_bar()
attribute_0x3 = ggplot(train, aes (x = attribute_0, fill=attribute_3)) + geom_bar()
attribute_1x0 = ggplot(train, aes (x = attribute_1, fill=attribute_0)) + geom_bar()
attribute_1x2 = ggplot(train, aes (x = attribute_1, fill=attribute_2)) + geom_bar()
attribute_1x3 = ggplot(train, aes (x = attribute_1, fill=attribute_3)) + geom_bar()
attribute_2x0 = ggplot(train, aes (x = attribute_2, fill=attribute_0)) + geom_bar()
attribute_2x1 = ggplot(train, aes (x = attribute_2, fill=attribute_1)) + geom_bar()
attribute_2x3 = ggplot(train, aes (x = attribute_2, fill=attribute_3)) + geom_bar()
grid.arrange(attribute_0x1, attribute_0x2, attribute_0x3)
grid.arrange(attribute_1x0, attribute_1x2, attribute_1x3)
grid.arrange(attribute_2x0, attribute_2x1, attribute_2x3)

cont_table = train %>% count(attribute_0, attribute_1, attribute_2, attribute_3)
cont_table
```





# Since there are only 5 combinations of attributes, combine them all into 1 column to identify patterns
```{r}
#first renaming attributes 0 and 1 to make it cleaner
train = train %>%  mutate(attribute_0 = fct_recode(attribute_0, "7" = "material_7", "5" = "material_5")) %>% mutate(attribute_1 = fct_recode(attribute_1, "8" = "material_8", "5" = "material_5", "6" = "material_6"))

combined_train = train %>% 
  unite("all_attributes", attribute_0, attribute_1, attribute_2, attribute_3, sep = "", remove = FALSE) %>% 
  mutate(all_attributes = as_factor(all_attributes))
head(combined_train)

#Matching to product codes
table = combined_train %>% count(product_code, all_attributes)
table

#Getting a clean table with all attributes
finaltable = combined_train %>% count(product_code, attribute_0, attribute_1, attribute_2, attribute_3)
productattributes = as.data.frame(finaltable)
finaltable = kable(productattributes, caption = "Product Attributes", align = "c", col.names = c("Product Code", "Attribute 0", "Attribute 1", "Attribute 2", "Attribute 3", "Number of samples"))
finaltable

```
