---
output:
  word_document: default
  html_document: default
  pdf_document: default
---

loading and measurement 17
```{r loading libraries}
library(tidyverse)
library(tidymodels)
library(GGally)
library(ggcorrplot)
library(gridExtra)
library(mice)
library(VIM)
library(naniar)
library(skimr)
library(UpSetR)

library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)

library(rpart) #for classification trees
library(rpart.plot) #for plotting trees
library(RColorBrewer) #better visualization of classification trees
library(rattle) #better visualization of classification trees

library(usemodels)
library(glmnet)


library(VIM) #visualizing missingness
library(ranger) #for random forests
library(randomForest) #also for random forests
library(caret)
library(skimr)
library(GGally)
library(gridExtra)
library(vip) #variable importance

library(xgboost)

library(nnet)

library(knitr)
library(kableExtra)

```

# Bringing in data set
```{r}
supersoaker <- read_csv("train.csv")

supersoaker = supersoaker %>% mutate(failure = as_factor(failure)) %>% mutate(product_code = as_factor(product_code)) %>% mutate(attribute_0 = as_factor(attribute_0)) %>% mutate(attribute_1 = as_factor(attribute_1)) %>% mutate(attribute_2 = as_factor(attribute_2)) %>% mutate (attribute_3 = as_factor(attribute_3)) %>% select(-id)

str(supersoaker)

```

# Imputing data
```{r}
imputation = mice(supersoaker, m=5, method = "pmm", printflag=FALSE)
complete_supersoaker = complete(imputation) 
summary(complete_supersoaker)

```

# Testing/training split
```{r}
data_split = initial_split(complete_supersoaker, prop=.7, strata=failure)
train = training(data_split)
test = testing(data_split)

```

# Setting up folds
```{r}
folds = vfold_cv(train, v=5)
```







# Linear Regression
```{r}
lm_recipe  = recipe(failure ~., train) %>%
  step_dummy (all_nominal(),-all_outcomes()) 

lm_model = logistic_reg(mode="classification") %>%
  set_engine("glm")

lm_workflow =  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(lm_recipe)

lm_fit  = fit(lm_workflow, train)

summary(lm_fit$fit$fit$fit)


predictions = predict(lm_fit, train, type= "prob") [2]
ROCRprediction = prediction(predictions, train$failure)
ROCRperformance = performance(ROCRprediction, "tpr", "fpr")
as.numeric(performance(ROCRprediction, "auc")@y.values)

opt.cut = function(perf, pred){
cut.ind = mapply(FUN=function(x, y, p){
d = (x - 0)^2 + (y-1)^2
ind = which(d == min(d))
c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
cutoff = p[[ind]])
}, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperformance, ROCRprediction))

table = table(train$failure,predictions > .2036858)

(table[1,1] + table[2,2])/nrow(train) #accuracy
table[2,2]/(table[2,2] + table[2,1]) #sensitivity
table[1,1]/(table[1,1] + table[1,2]) #specificity

```


# Lasso regression - first testing the significant variables identified in the linear regression model 
```{r}
lasso_recipe = recipe(failure ~ loading + measurement_2 + measurement_4 + measurement_6 + measurement_7 + measurement_8, train) %>% 
  step_dummy(all_nominal(),-all_outcomes())


lasso_model = logistic_reg(penalty = tune(), mixture = 1) %>% 
set_mode("classification") %>% 
set_engine("glmnet")

lasso_workflow =  workflow() %>% 
  add_recipe(lasso_recipe) %>% 
  add_model(lasso_model)

glmnet_grid = grid_regular(penalty(), levels = 200)

glmnet_tune <- 
  tune_grid(lasso_workflow, resamples = folds, 
            grid = glmnet_grid, metrics = metric_set(mn_log_loss))
```

Visualizing tuning parameters
```{r}
glmnet_tune %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  theme(legend.position = "none")
```

Finding best mn_log_loss
```{r}
best_mnlog = glmnet_tune %>%
  select_best(metric="mn_log_loss")
best_mnlog
```

Finalizing fit
```{r}
final_lasso = lasso_workflow %>% finalize_workflow(best_mnlog)
lasso_fit = fit(final_lasso, train)

tidy(lasso_fit)
```

Running predictions on training set
```{r}
predictions = predict(lasso_fit, train, type= "prob")[2]
ROCRpred = prediction(predictions, train$failure)
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```

Finding cutoff to balance of sensitivity and specificity
```{r}
opt.cut = function(perf, pred){
cut.ind = mapply(FUN=function(x, y, p){
d = (x - 0)^2 + (y-1)^2
ind = which(d == min(d))
c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
cutoff = p[[ind]])
}, perf@x.values, perf@y.values, pred@cutoffs)
}
best_cutoff = opt.cut(ROCRperf, ROCRpred)
```

Seeing cutoff score accuracy
```{r}
lm_table_train = table(train$failure, predictions > best_cutoff[3])
accuracy_lm_table_train = (lm_table_train[1,1] + lm_table_train[2,2])/nrow(train)
accuracy_lm_table_train
```

Predictions on the testing set
```{r}
predictions_test = predict(lasso_fit, test, type= "prob")[2]
ROCRpred = prediction(predictions_test, test$failure)
ROCRperf = performance(ROCRpred, "tpr", "fpr")
```

```{r}
lm_table_test = table(test$failure, predictions_test > best_cutoff)
accuracy_lm_table_test = (lm_table_test[1,1] + lm_table_test[2,2])/nrow(test)
accuracy_lm_table_test
```




# Typical Lasso regression, using all variables

```{r}
lasso_recipe_all = recipe(failure ~., train) %>% 
  step_unknown(all_nominal_predictors()) %>% #added this step since the final testing set has different attributes and product codes than found in the training set
  step_dummy(all_nominal(),-all_outcomes())


lasso_model_all = logistic_reg(penalty = tune(), mixture = 1) %>% 
set_mode("classification") %>% 
set_engine("glmnet")

lasso_workflow_all =  workflow() %>% 
  add_recipe(lasso_recipe_all) %>% 
  add_model(lasso_model_all)

glmnet_grid_all = grid_regular(penalty(), levels = 100)

glmnet_tune_all <- 
  tune_grid(lasso_workflow_all, resamples = folds, 
            grid = glmnet_grid, metrics = metric_set(mn_log_loss))
```

```{r}
best_mnlog_all = glmnet_tune_all %>%
  select_best(metric="mn_log_loss")
```

```{r}
final_lasso_all = lasso_workflow_all %>% finalize_workflow(best_mnlog_all)
lasso_fit_all = fit(final_lasso_all, train)

tidy(lasso_fit_all)
```

```{r}
predictions_all = predict(lasso_fit_all, train, type= "prob")[2]
ROCRpred_all = prediction(predictions_all, train$failure)
ROCRperf_all = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf_all, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```

```{r}
opt.cut = function(perf, pred){
cut.ind = mapply(FUN=function(x, y, p){
d = (x - 0)^2 + (y-1)^2
ind = which(d == min(d))
c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
cutoff = p[[ind]])
}, perf@x.values, perf@y.values, pred@cutoffs)
}
best_cutoff = opt.cut(ROCRperf_all, ROCRpred_all)
```

```{r}
lm_table_train_all = table(train$failure, predictions_all > best_cutoff[3])
accuracy_lm_table_train_all = (lm_table_train_all[1,1] + lm_table_train_all[2,2])/nrow(train)
accuracy_lm_table_train_all

```


```{r}
predictions_test_all = predict(lasso_fit_all, test, type= "prob")[2]
ROCRpred_alltest = prediction(predictions_test_all, test$failure)
ROCRperf_alltest = performance(ROCRpred_alltest, "tpr", "fpr")

lm_table_test_all = table(test$failure, predictions_test_all > best_cutoff[3])
accuracy_lm_table_test_all = (lm_table_test_all[1,1] + lm_table_train_all[2,2])/nrow(test)
accuracy_lm_table_test_all
```
Better accuracy with all predictors - Loading, Measurements: 2, 4, 8, 11, 13, 17, Product code, Attributes: 0, 2 used in the model 







# Classification tree
```{r}
tree_recipe = recipe(failure ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree() %>% 
 set_engine("rpart", model = TRUE) %>% 
set_mode("classification")

tree_workflow =  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(tree_recipe)

tree_fit = fit(tree_workflow, train)

tree_fit %>%
  pull_workflow_fit() %>%
  pluck("fit") 

tree = tree_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(tree)
```
End tree does not have any branches - only categorizes "No" 









# Random forest

```{r}
rf_recipe = recipe(failure ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

rf_model = rand_forest() %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

rf_workflow =  workflow() %>% 
 add_model(rf_model) %>% 
 add_recipe(rf_recipe)

rf_fit = fit(rf_workflow, train)
rf_fit

pred = predict(rf_fit, train)
confusionMatrix(pred$.pred_class,train$failure, positive="Yes")

rf_fit %>%
extract_fit_parsnip() %>%
vip(geom = "point")
```

Tuning mtry and min_n
```{r}
rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

rf_recipe = recipe(failure ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_workflow =  workflow() %>% 
 add_model(rf_model) %>% 
 add_recipe(rf_recipe)

```

Tuning cont. 
```{r}
rf_residuals = tune_grid(rf_workflow, resamples = folds, grid = 20)

rf_residuals %>%
collect_metrics() %>%
filter(.metric == "accuracy") %>%
select(mean, min_n, mtry) %>%
pivot_longer(min_n:mtry,
values_to = "value",
 names_to = "parameter") %>%
ggplot(aes(value, mean, color = parameter)) +
geom_point(show.legend = FALSE) +
facet_wrap(~parameter, scales = "free_x") +
labs(x = NULL, y = "Accuracy")
```

Setting tuning parameters based on graph above
```{r}
rf_grid = grid_regular(
  mtry(range = c(0, 15)), 
  min_n(range = c(50, 150)),
  levels = 5)

```

Creating another graph of accuracy with selected parameters
```{r}
tuned_residuals = tune_grid(rf_workflow, resamples = folds, grid = rf_grid)
```
```{r}
tuned_residuals %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

Selecting best parameters
```{r}
best_rf = select_best(tuned_residuals, metric="accuracy")

final_rf = finalize_workflow(
  rf_workflow,
  best_rf
)

final_rf
```

Finalizing and saving fit
```{r}
final_rf_fit = fit(final_rf, train)
saveRDS(final_rf_fit, "final_rf_fit.rds")
```

```{r}
final_rf_fit = readRDS("final_rf_fit.rds") 
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```


Running predictions on training data
```{r}
trainpredrf = predict(final_rf_fit, train)
cm_rf_train = confusionMatrix(trainpredrf$.pred_class, train$failure, 
                positive = "Yes")
accuracy_rf_train = cm_rf_train$overall["Accuracy"]
accuracy_rf_train
```


Running predictions on testing data
```{r}
testpredrf = predict(final_rf_fit, test)
cm_rf_test = confusionMatrix(testpredrf$.pred_class, test$failure, 
                positive = "Yes")
accuracy_rf_test = cm_rf_test$overall["Accuracy"]
accuracy_rf_test
```











# XGBoost

Getting starting code
```{r}
#use_xgboost (failure ~., data = train)
```

```{r}

start_time = Sys.time()

xgboost_recipe <- 
  recipe(formula = failure ~ ., data = train) %>% 
  step_dummy(all_nominal(), -all_outcomes(), one_hot=TRUE) %>%
  step_zv(all_predictors())
  

xgboost_spec <- 
  boost_tree(trees = tune(), min_n = tune(), tree_depth = tune(), learn_rate = tune(),
    loss_reduction = tune(), sample_size = tune()) %>% 
  set_mode("classification") %>% 
  set_engine("xgboost") 

xgboost_workflow <- 
  workflow() %>% 
  add_recipe(xgboost_recipe) %>% 
  add_model(xgboost_spec) 

xgboost_tune <-
  tune_grid(xgboost_workflow, resamples = folds, grid = 25)

end_time = Sys.time()
end_time - start_time 

```

Best parameters
```{r}
best_xgboost = select_best(xgboost_tune, metric = "accuracy")
xgboost_finalwf= finalize_workflow(xgboost_workflow, best_xgboost)

```

Finalizing fit
```{r}
xgboost_finalfit = fit(xgboost_finalwf, train)
saveRDS(xgboost_finalfit, "xgboostfit")

```

Running predictions on training data
```{r}
xgboost_finalfit = readRDS("xgboostfit")
trainpredxg = predict(xgboost_finalfit, train)
cm_xgb_train = confusionMatrix(trainpredxg$.pred_class,train$failure, positive="Yes")
accuracy_xgb_train = cm_xgb_train$overall["Accuracy"]
accuracy_xgb_train

```

Running predictions on testing data
```{r}
testpredxg = predict(xgboost_finalfit, test)
cm_xgb_test = confusionMatrix(testpredxg$.pred_class,test$failure, positive="Yes")
accuracy_xgb_test = cm_xgb_test$overall["Accuracy"]
accuracy_xgb_test
```



# Neural Network
```{r}
nn_start = Sys.time()

nn_recipe = recipe(failure ~., train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%  
  step_dummy(all_nominal(), -all_outcomes())


nn_model = mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_mode("classification") %>% 
  set_engine("nnet", verbose = 0)

nn_workflow = workflow() %>% 
  add_model(nn_model) %>% 
  add_recipe(nn_recipe)

nn_tune = tune_grid(nn_workflow, resamples = folds, grid = 25)

nn_end = Sys.time()

nn_start-nn_end

```

Looking at the tuning parameters
```{r}
nn_tune %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, hidden_units, penalty, epochs) %>%
  pivot_longer(hidden_units:epochs, values_to = "value", names_to = "parameter") %>%
  ggplot(aes(value, mean, color = parameter)) + geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") + labs(x = NULL, y = "Accuracy")

```

Fitting the best parameters
```{r}
best_nn = select_best(nn_tune, metric = "accuracy")

final_nn = finalize_workflow(nn_workflow, best_nn)
final_nn
final_nn_fit = fit(final_nn, train)

saveRDS(final_nn_fit, "nnfit")

```

Making predictions on training data
```{r}

final_nn_fit = readRDS("nnfit")
nn_predtrain = predict(final_nn_fit, train)
cm_nn_train = confusionMatrix(nn_predtrain$.pred_class,train$failure, positive="Yes")
accuracy_nn_train = cm_nn_train$overall["Accuracy"]
accuracy_nn_train
```

Making predictions on testing data
```{r}
nn_predtest = predict(final_nn_fit, test)
cm_nn_test = confusionMatrix(nn_predtest$.pred_class,test$failure, positive="Yes")
accuracy_nn_test = cm_nn_test$overall["Accuracy"]
accuracy_nn_test
```


# Running models on final test set

## Importing and imputing the final test set
```{r}
new_test <- read_csv("test.csv")

new_test = new_test %>% mutate(product_code = as_factor(product_code)) %>% mutate(attribute_0 = as_factor(attribute_0)) %>% mutate(attribute_1 = as_factor(attribute_1)) %>% mutate(attribute_2 = as_factor(attribute_2)) %>% mutate (attribute_3 = as_factor(attribute_3))

```

```{r}
imputation_final = mice(new_test, m=5, method = "pmm", printflag=FALSE)
final_test = complete(imputation_final) 
summary(final_test)

```


# Lasso model seemed most promising - only one that categorized more than 20 instances as failed (greater variance from the No Information Rate)
```{r}
predictions_final = predict(lasso_fit_all, final_test, type= "prob")[2]
  

predictions_final = predictions_final %>% mutate(FAILURE = if_else(.pred_Yes > best_cutoff[3], "Yes", "No"))

final_lm = data.frame(final_test$id, predictions_final$FAILURE)

final_lm = final_lm %>% mutate(FAILURE = predictions_final$FAILURE) %>% mutate(ID = final_test$id) %>% select(ID, FAILURE)
head(final_lm)

write_csv(final_lm, "final_lm.csv")
```


## Running Random Forests (technically had the highest accuracy on the test/train data)

```{r}
final_rf_pred = predict(final_rf_fit, final_test)
final_rf = final_test %>% mutate(FAILURE = final_rf_pred$.pred_class) %>% mutate(ID = id) %>% select(ID, FAILURE)

write_csv(final_rf, "final_rf.csv")

```



## Also want to see how XGBoost runs 

```{r}
final_testpredxg = predict(xgboost_finalfit, final_test)

final = final_test %>% mutate(FAILURE = final_testpredxg$.pred_class) %>% mutate(ID = id) %>% select(ID, FAILURE)


write_csv(final, "final.csv")

```


After submission, the Lasso model performed the best. 






# Building visualizations for the presentation

## Showing accuracy tests

```{r}
# Lasso Model
lasso_df = data.frame(Dataset = c("Training", "Testing"), Accuracy = c(accuracy_lm_table_train_all, accuracy_lm_table_test_all))
lasso_table = kable(lasso_df, caption = "Lasso Regression", align = "c")
lasso_table

# Random Forest Model
rf_df = data.frame(Dataset = c("Training", "Testing"), Accuracy = c(accuracy_rf_train, accuracy_rf_test))
rf_table = kable(rf_df, caption = "Random Forest", align = "c")
rf_table

# XGBoost 
xgb_df = data.frame(Dataset = c("Training", "Testing"), Accuracy = c(accuracy_xgb_train, accuracy_xgb_test))
xgb_table = kable(xgb_df, caption = "XGBoost", align = "c")
xgb_table

# Neural Network
nn_df = data.frame(Dataset = c("Training", "Testing"), Accuracy = c(accuracy_nn_train, accuracy_nn_test))
nn_table = kable(nn_df, caption = "Neural Network", align = "c")
nn_table

```

## Confusion matrixes 

```{r}

#Lasso
lm_table_test_all

#Random Forest
cm_rf_test$table

#XGBoost
cm_xgb_test$table


#Neural Network
cm_nn_test$table

```

